\section{Hypothesis Ledger}

This ledger records live hypotheses. It is deliberately conservative: hypotheses may be beautiful, but they must become testable.

\begin{longtable}{p{0.08\textwidth}p{0.48\textwidth}p{0.16\textwidth}p{0.20\textwidth}}
\toprule
\textbf{ID} & \textbf{Hypothesis} & \textbf{Current Level} & \textbf{First Test} \\
\midrule
\endhead
H0 & The system is primarily non-linguistic: emblematic, administrative, cultic, or owner/lineage marking. & Testable & Compare brevity, repetition, positional rigidity, and object-class specificity against non-linguistic baselines. \\
H1 & The system is restricted writing: linguistic or para-linguistic formulae, not full prose. & Testable & Look for stable positional classes, formula slots, and controlled alternations across sites/media. \\
H2 & The system is logo-syllabic or logo-phonetic with many allographs. & Speculative & Test whether sign inventory can be compressed into a smaller functional inventory without losing positional structure. \\
H3 & A Dravidian substrate explains some sign clusters or terminal elements. & Speculative & Search for suffix-like positional behavior and compare only after corpus classes are established. \\
H4 & Indo-Aryan or Indo-Iranian readings explain some sign clusters. & Speculative & Require chronology-sensitive linguistic forms and predictions beyond isolated lexical resemblance. \\
H5 & The corpus reflects a multilingual administrative ecology using language-neutral signs plus limited phonetic elements. & Speculative & Test whether regional/site differences cluster by artifact function more than by putative language. \\
H6 & Numeral-like signs and commodities are the easiest first partial decipherment target. & Suggestive & Isolate stroke/numeral candidates, following-sign distributions, and correlations with object types, weights, or measures. \\
\bottomrule
\end{longtable}

\subsection{First Decision}

The first serious analysis should not be a reading. It should be a positional grammar:

\begin{enumerate}
  \item Normalize directionality.
  \item Extract start, end, and medial sign distributions.
  \item Split by artifact type and site.
  \item Identify signs whose positional entropy is unusually low.
  \item Use those signs to define candidate functional classes.
\end{enumerate}

