\section{Source Register}

This register tracks sources we may use and how much evidentiary weight they should carry.

\begin{longtable}{p{0.25\textwidth}p{0.20\textwidth}p{0.16\textwidth}p{0.30\textwidth}}
\toprule
\textbf{Source} & \textbf{Use} & \textbf{Status} & \textbf{Notes} \\
\midrule
\endhead
Local \texttt{data/ivs\_corpus\_cleaned.csv} & Initial working corpus & Active, provenance to verify & 5,679 rows; excludes claimed Sanskrit readings/translations. \\
Local \texttt{data/ivs\_corpus.csv} & Hypothesis archive only & Restricted & Includes proposed Sanskrit readings/translations; not baseline data. \\
Yajnadevam \texttt{lipi} repository \citep{yajnadevamLipi} & Public source match for local CSV structure & Active provenance lead & Exposes an \texttt{inscriptions.csv} stream whose fields and sample rows align with the local corpus. \\
Fuls/ICIT documentation \citep{fulsICIT} & Directionality and text-code conventions & Active provenance lead & Documents the interactive corpus and its R/L text-code search conventions. \\
Fuls, ``Positional Analysis of Indus Signs'' \citep{fuls2013positional} & Positional method and comparison panels & Tracked local PDF & Stored at \texttt{references/pdfs/fuls/Fuls\_2013\_Positional\_Analysis\_of\_Indus\_Signs.pdf}. PDF page 10 contains a panel for \texttt{692}, closing the missing visual side for the \texttt{692}/\texttt{920} Tier 1 pair. \\
Fuls, \emph{Corpus of Indus Inscriptions} \citep{fuls2022corpus} & ICIT corpus and sign-list reference & Tracked local PDF & Stored at \texttt{references/pdfs/fuls/Fuls\_2022\_Corpus\_of\_Indus\_Inscriptions.pdf}. PDF page 21 contains a high-number sign-list panel covering ICIT \texttt{703}--\texttt{958}, including \texttt{705}, \texttt{706}, \texttt{817}, \texttt{820}, \texttt{861}, and \texttt{920}. \\
Harappa catalog review \citep{harappaCatalog2026} & Public description of Fuls' sign catalog & Contextual reference & Confirms that the catalog is visually detailed and designed as a sign-reference tool, but does not provide a machine-readable target-code image table. \\
Deutsche Digitale Bibliothek record \citep{ddbFulsCatalog2026} & Bibliographic access to Fuls' catalog & Contextual reference & Confirms publication details for the 560-page illustrated catalog, while noting that only the table of contents is digitally open. \\
Local Fuls 2023 SearchInside PDF sample & Supplementary ICIT visual mapping & Tracked local PDF & Stored at \texttt{references/pdfs/fuls/Fuls\_2023\_Catalog\_of\_Indus\_Signs\_SearchInside.pdf}. The sample independently shows additional catalog visuals, including \texttt{820} and \texttt{920}. \\
Daggumati and Revesz \citep{daggumati2021allographs} & Allograph-review method & Active methodological reference & Provides a cautious framework for testing redundancies, mirroring, directionality, space-saving, catalog errors, and image-based allograph claims. \\
Mahadevan, \emph{The Indus Script: Texts, Concordance and Tables} \citep{mahadevan1977} & Standard concordance reference & Foundational; tracked local PDF & Stored at \texttt{references/pdfs/mahadevan/Mahadevan\_1977\_The\_Indus\_Script\_Texts\_Concordance\_and\_Tables.pdf}. Often cited as M77; sign counts and text lines differ depending on filtering. \\
Joshi and Parpola, \emph{Corpus of Indus Seals and Inscriptions}, Vol. 1 \citep{joshiParpola1987cisi1} & Artifact-image authority & Foundational; tracked local PDF & Stored at \texttt{references/pdfs/cisi/Joshi\_Parpola\_1987\_CISI\_Vol1\_Collections\_in\_India.pdf}. Useful later for seal/inscription image checks once ICIT/CISI/page links are built. \\
Marshall, \emph{Mohenjo-daro and the Indus Civilization}, Vol. II \citep{marshall1931mohenjodaro2} & Excavation context and early text discussion & Tracked local PDF & Stored at \texttt{references/pdfs/marshall/Marshall\_1931\_Mohenjo\_Daro\_and\_the\_Indus\_Civilization\_Vol2.pdf}. Potentially useful for archaeological context and older readings, not a baseline sign-list authority. \\
CISI remaining volumes & Corpus/reference tradition & Candidate future sources & Additional volumes may be needed for Pakistan and later-discovered materials. \\
mayig \texttt{indus-valley-script-corpus} & Open digital CISI-oriented corpus & Candidate comparative corpus & WIP digitization of CISI in JSON; records graphemes left-to-right while assuming right-to-left reading for the script. \\
Yadav et al. \citep{yadav2010plosone} & N-gram/statistical baseline & Active methodological reference & Reports Zipf-Mandelbrot behavior, positional asymmetry, strong bigram correlations, and Markov modeling. \\
Rao et al. \citep{rao2009science} & Conditional entropy comparison & Active methodological reference & Argues Indus conditional entropy resembles linguistic systems more than tested non-linguistic controls. \\
Rao et al. \citep{rao2009pnas} & Markov modeling & Active methodological reference & Uses positional statistics and Markov methods for syntax and missing-sign prediction. \\
Farmer, Sproat, and Witzel \citep{farmer2004collapse} & Non-linguistic critique & Required adversarial model & Argues against a linguistic-script thesis and stresses brevity, lack of long texts, and non-linguistic symbol-system parallels. \\
Sinha et al. \citep{sinha2010network} & Network analysis & Candidate methodological reference & Reports network patterns suggestive of syntactic organization. \\
Nair \citep{nair2026progress} & Recent synthetic non-linguistic baseline & Provisional & New preprint; useful as an emerging baseline design, not yet treated as settled. \\
\bottomrule
\end{longtable}

\subsection{Immediate Source Tasks}

\begin{enumerate}
  \item Identify the provenance of the two local CSV files.
  \item Determine the sign numbering system used locally.
  \item Build concordance crosswalks where possible: local code, Mahadevan, Parpola, Wells.
  \item Build a local PDF page index for the curated reference library.
  \item Link ICIT/CISI identifiers to artifact-image pages in the local CISI volume.
  \item Decide whether to import the open JSON corpus as a second corpus layer.
\end{enumerate}
